Mechanized Reasoning about a Capability Machine

### 2019-10-16 prisc-2020/paper.tex
one more comment
changes to section 3

\todo{Alix: Add a sentence to say that we artificially add expressions to fit into the framework, but the operational semantics is equivalent to the standard one}
Strictly speaking, a machine does not have expressions. Rather the small step semantics are defined over machine configurations, and we want the program logic to illustrate that. Like in \cite{Jensen:2013}, we define a notion of points-to for registers. Since a capability machine replaces pointers with capabilities, we will also replace the conventional points-to for memory with a points-to assertion with a permission attached to it: 
$$ r \mapsto_r w \hspace{1cm} a \mapsto_a[p] w $$

@@

\subsection{Definition of the Logical Relation}
The logical relation is a unary relation that relates a word to  \textit{capability safety}. Figure \ref{fig:relation} describes the Iris implementation of the logical relation defined by Skorstensgaard et. al. \cite{skorstengaardESOP18}, in which the authors define a capability machine with so-called \emph{local} capabilities, which imposes certain restrictions on where to store such capabilities. The value relation $\mathcal{V}$ is an Iris relation of type $\world \to \word \to \emph{iProp}$, where \world{} is a collection of state transition systems used to reason about local states\todo{Alix: explain local as in local to a function call}, and a Word can be an integer of a capability. We then define the register relation $\mathcal{R}$ as the validity of each word that the register state maps to.

The expression relation should capture what it means for programs to be capability safe. However, as previously discussed, programs in a machine lie in memory, pointed to by the program counter. Strictly speaking, there are no "expressions" in such a machine language. Instead, the expression relation relates words to the notion of program capability safety, where the word in question is pointed to by the program counter.

\subsection{The Context}
\todo{Everything ghost might be too much of a Iris'sm ?}
In the expression relation, the \textsf{context} contains the ghost resources necessary to run a program pointed to by the program counter. When the program has finished running, ownership of these resources is given back. To start out, the running program gets full ownership of all machine registers. A ghost register will be represented by the full ownership of the map from RegName to Word. We write 
$$ r \mapsto_r w$$
to indicate that registers r currently contains the word w. The context will contain such a statement for each register in the register map.

@@

One way to enforce this, is to define the value relation of a capability as an invariant containing the resources for its range of authority, which at the same time guarantees that all the words that these words point to are also valid. However, as the name indicates, an invariant can never change. This is not a problem when the only variant is the current state of a location. In this case however, the value relation is indexed over a particular \world. This World will capture the behaviour of local state. The validity of an executable capability will consider any possible future world. This becomes crucial when showing safety of programs that depend on well bracketed control flow.

\subsection{The World}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.

>>>>

\todo{Alix: Add a sentence to say that we artificially add expressions to fit into the framework, but the operational semantics is equivalent to the standard one}
Strictly speaking, a machine does not have expressions. Rather the small step semantics are defined over machine configurations, and we want the program logic to illustrate that. Like in \cite{Jensen:2013}, we define a notion of points-to for registers. Since a capability machine replaces pointers with capabilities, we will also replace the conventional points-to for memory with a points-to assertion with a permission attached to it: 
$$ r \mapsto_r w \hspace{1cm} a \mapsto_a[p] w $$
\todo{Alix: give a quick introduction to permissions, and says for instance that a RO permission does not allow to update the ressource}

@@

\subsection{Definition of the Logical Relation}
The logical relation is a unary relation that relates a word to a notion of \textit{capability safety}. The value relation $\mathcal{V}$ is an Iris relation of type $World \to Word \to iProp$, where the World is a collection of state transition systems used to reason about local state\todo{Alix: explain local as in local to a function call}, and a Word can be an integer or a capability. We then define the register relation $\mathcal{R}$ as the validity of each word that the register state maps to.

The expression relation should capture what it means for programs to be capability safe. However, as previously discussed, programs in a machine lie in memory, pointed to by the program counter. The expression relation relates Words to the notion of program capability safety, where the word in question will be the current state of the program.

\subsection{The Context}
\todo{Everything ghost might be too much of a Iris'sm ?}
In the expression relation, the \textsf{context} contains the ghost resources necessary to run a program pointed to by the program counter. When the program has finished running, ownership of these resources is given back.

To start out, the running program gets full ownership of all machine registers. A ghost register will be represented by the full ownership of the map from RegName to Word. The context will contain a points-to assertion for each register in the register map.

@@

One way to enforce this, is to define the value relation of a capability as an invariant containing the resources for its range of authority, which at the same time guarantees that all the words that these words point to are also valid. The memory resource will then be given through the $\mathcal{R}$ relation.

However, as the name indicates, an invariant can never change. This is not a problem when the only variant is the current state of a location. In this case however, the value relation is parametrized over a particular World. The World captures the behaviour of local state. The validity of an executable capability will consider any possible future world. This becomes crucial when showing safety of programs that depend on well bracketed control flow. We do not want the invariant to depend on any one World.

Instead, we use Iris \textit{saved predicates}. The context contains the full authority over a map from locations to saved predicate names, asserting the resource for each location in the map, as well as the saved predicate applied to the current World and the word the location maps to. The map itself is abstracted by an existential quantifier. The Iris proposition rel($a,p,\mathcal{V}$) then states that the location a is included in that map, the points to relation has permission p, and the saved predicate is the value relation.
\newpage
%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
